Bounded Model Checking for Embedded Systems \\ (Position Paper)

### 2016-01-04 paper.tex
comments regarding the whole paper

Generally, embedded computer systems perform dedicated functions with high degree of reliability. They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}. Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems, which demand short development cycles and high-level of reliability~\cite{leeCPS}. As a consequence, human life has also become more and more dependent on the services provided by this type of system; in particular, their success depends both on the provided service relevance and quality.

Indeed, many current embedded systems become interesting solutions only if they can reliably
performs their target tasks, such as unmanned aerial vehicles (UAVs) and medical monitoring systems. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable. In addition, wrong diagnosis or condition-evaluation reports also have the potential to compromise patients' health, with serious consequences.

A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy. Additionally, the increasing computational power and decreasing size and cost common to computer processors is enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.

@@

State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages. In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness. For instance, such aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers.
%In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification. These aspects are of extreme relevance in micro-grids to ensure reliability, which is a key issue for (smart) cities and consumers.

@@

A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}. In the base-case, we aim to check that $\phi$ holds in all states reachable from an initial state within $k$ steps (we assume that $k \geq 0$) and in the induction-step, we aim to check that whenever $\phi$ holds in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next state $s_{k+1}$ of the system. In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints specified during the system design phase~\cite{EenS03}. Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}. Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a new algorithm for proving correctness of (a large set of) C programs by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}. However, there is still little evidence in the associated literature that model checking embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.

%--------------------------------------
\subsection{Incorporating System Knowledge to Verification}
%--------------------------------------

@@

Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, Husamma {\it et al.} proposed a verification tool for digital systems to aid enginners to check for overflow, limit cycle, error, timing, stability, and minimum phase in digital systems, considering finite word length (FWL) effects~\cite{dsv_spin2015}.

@@

\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++98 programs~\cite{ECBS13}. \textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques to prune the state-space exploration~\cite{morse15}. These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC). \textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15}. \textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}. Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems) as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.

@@

This paper presented the main challenges to verify design correctness in embedded systems. In particular, it emphasizes that stringent constraints imposed by the hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification. Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
extends BMC-based approaches from falsification to verification, can be applied to formally verify the correctness and the timeliness of embedded systems. Given that the complexity of software has increased significantly in embedded products, there are still some (recent) advances to stress and cover exhaustively the state space of the system, in order to verify low-level properties that have to meet the application's deadline, access the memory region, handle concurrency, and control the hardware registers. As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded systems implementation; special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification.

>>>>

Generally, embedded computer systems perform dedicated functions with high degree of reliability. They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}. Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems, which demand short development cycles and high-level of reliability~\cite{leeCPS}. As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success depends both on the provided service relevance and quality.

Indeed, many current embedded systems become interesting solutions only if they can reliably
perform their target tasks, such as unmanned aerial vehicles (UAVs) and medical monitoring systems. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable. In addition, wrong diagnosis or condition-evaluation reports also have the potential to compromise patients' health, with serious consequences.

A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy. Additionally, the increasing computational power and decreasing size and cost, which are common to computer processors, are enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.

@@

State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages. In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness. For instance, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third is essential in systems that implement scientific models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions.
%In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification. These aspects are of extreme relevance in micro-grids to ensure reliability, which is a key issue for (smart) cities and consumers.

@@

A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}. In the base-case, we aim to check that $\phi$ holds in all states reachable from an initial state within $k$ steps (we assume that $k \geq 0$) and in the induction-step, we aim to check that whenever $\phi$ holds in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next state $s_{k+1}$ of the system. In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints specified during the system design phase~\cite{EenS03}. Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}. Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a new algorithm for proving correctness of (a large set of) C programs by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}. However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.

%--------------------------------------
\subsection{Incorporating System Knowledge to Verification Procedures}
%--------------------------------------

@@

Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check for overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects. DSverifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions. Ultimately, if some requirements are not met with a given configuration (realization structure and hardware restrictions), the error reports provided DsVerifier may suggest another setup. Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAVs.

@@

\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++98 programs~\cite{ECBS13}. \textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques to prune the state-space exploration~\cite{morse15}. These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in, standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC). \textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15}. \textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} \textcolor{red}{depois mencionar explicitamente UAVS} and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}. Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems) as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.

@@

This paper presented the main challenges to verify design correctness in embedded systems. In particular, it emphasizes that stringent constraints imposed by the hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification. Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
extends BMC-based approaches from falsification to verification, can be applied to formally verify the correctness and the timeliness of embedded systems. Given that the complexity of software has increased significantly in embedded products, there are still some (recent) advances to stress and cover exhaustively the state space of the system, in order to verify low-level properties that have to meet the application's deadline, access the memory region, handle concurrency, and control the hardware registers \textcolor{red}{falar sobre a incorporacao e conhecimento sobre o sistema}. As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded systems implementation; special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification.
